---
title: "Project_Solar_Energy"
author: "Nguyen Nguyen"
date: "April 1, 2019"
output: rmarkdown::github_document
---

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
#install.packages('corrplot')
library(corrplot)
library(stringr)
#install.packages('BBmisc') #correlation plot
library(BBmisc)
library(ggplot2)
#install.packages('ggfortify')  #mapping biplot
library(ggfortify)
#install.packages('tidyverse')
library(tidyverse)
#install.packages('usmap') #us map by state or county
library(usmap)
#install.packages('factoextra')
library(factoextra)
library(mclust)
library(sem)
library(semPlot)
#install.packages('rgl') #3D plotting
library(rgl)
library(reshape2)
```


**County-level aggregation**
#Using SUM 
```{r message=FALSE, warning=FALSE}
county_solar_potential <- read_csv("C:/Users/nguye/Documents/TTU/6350/Group Project/county_solar_potential.csv")
county_solar_potential <-county_solar_potential[,-1]
capacity_potential_lmi <- county_solar_potential %>%
  select(state_name:county_name,area_km2,very_low_mf_own_mw:mod_sf_rent_mw) 
# Correlation matrix
cormatrix <- round(cor(capacity_potential_lmi[,4:15]),4)
```

```{r}
head(capacity_potential_lmi)
```

#Check NA
```{r}
colSums(is.na(capacity_potential_lmi))
```

#Observations with low values, assigned as 0
```{r}
capacity_potential_lmi$COUNT <- rowSums(abs(capacity_potential_lmi[,4:15]) < 0.000000000000001)
sum(capacity_potential_lmi$COUNT > 5)
```

#counties with nearly more than >= 50% variables of zero values 
```{r}
capacity_potential_lmi[capacity_potential_lmi$COUNT>5,]
```

#remove counties with from more than >= 50% variables of zero values 
```{r}
capacity_potential_lmi_clean <- capacity_potential_lmi[capacity_potential_lmi$COUNT<6,]
```

#Scatter plot
```{r,dpi=1200,fig.height = 12, fig.width = 12, fig.align = "center"}
pairs(scale(capacity_potential_lmi_clean[,4:15]))
```

#Extract numerical parts - **all counties**
```{r}
num_vabs_cp_lmi <- as.data.frame(capacity_potential_lmi_clean[,4:15])

```

#Correlation plot
```{r}
corr_mat = round(cor(num_vabs_cp_lmi),2)
corrplot(corr_mat)
```


#Combine names of states and counties
```{r}
a <- abbreviate(capacity_potential_lmi_clean$state_name)
b <- abbreviate(capacity_potential_lmi_clean$county_name)
rownames(num_vabs_cp_lmi) <- str_c(a,b,sep='_')
```

*Outliers*
```{r}
#Observing very_low_mf_rent v. very_low_sf_own
plot(num_vabs_cp_lmi$very_low_mf_rent_mw,num_vabs_cp_lmi$very_low_sf_own_mw,type='n',xlab='very_low_mf_rent',ylab='very_low_sf_own')
text(num_vabs_cp_lmi$very_low_mf_rent_mw,num_vabs_cp_lmi$very_low_sf_own_mw,labels=rownames(num_vabs_cp_lmi))
#Observing low_mf_rent v. very_low_sf_rent
plot(num_vabs_cp_lmi$low_mf_rent_mw,num_vabs_cp_lmi$very_low_sf_rent_mw,type='n',xlab='low_mf_rent',ylab='very_low_sf_rent')
text(num_vabs_cp_lmi$low_mf_rent_mw,num_vabs_cp_lmi$very_low_sf_rent_mw,labels=rownames(num_vabs_cp_lmi))
```
Based on above 2 plots, potential outliers: California_LosAngeles. 


**Standardize the dataset**
```{r}
num_vabs_cp_lmi.s <- normalize(num_vabs_cp_lmi, method = "standardize", range = c(0, 1), margin = 2, on.constant = "quiet")
```


*Sun Belt states*
```{r message=FALSE, warning=FALSE}
sunbelt <- c('Florida','Georgia','South Carolina','Alabama','Mississippi', 'Louisiana','Texas','New Mexico','Arizona','Nevada','California')
sunbelt %in% capacity_potential_lmi_clean$state_name
capacity_potential_lmi_clean$sunbelt <- 'Other'
capacity_potential_lmi_clean$sunbelt[capacity_potential_lmi_clean$state_name %in% sunbelt] <- 'SunBelt'
```



**Principle Component Analysis for County-level**
#Assumption: dataset is multivariate normality.
# Principal Component Analysis - County level
```{r}
lmi_pca <- princomp(num_vabs_cp_lmi.s, cor=T)
summary(lmi_pca,loadings=T)
```

```{r}
head(lmi_pca$scores)
```


```{r,dpi=1200,fig.height = 6, fig.width = 6, fig.align = "center"}
biplot(lmi_pca) 
```


```{r}
num_vabs_cp_lmi.s$sunbelt <- capacity_potential_lmi_clean$sunbelt
```


```{r,dpi=1200,fig.height = 6, fig.width = 6, fig.align = "center"}
autoplot(princomp(num_vabs_cp_lmi[,-13],cor=T),loadings=TRUE,loadings.label=TRUE,loadings.label.size=5,loadings.label.colour='black',data=num_vabs_cp_lmi.s,colour='sunbelt',loadings.colour='black',size=5)+
  theme(legend.position="bottom",legend.title=element_text(size=15),legend.text=element_text(size=14))

```


**State level** 
#Extract numerical parts - *states only*
```{r}
capacity_potential_lmi_state <- capacity_potential_lmi_clean %>%
  group_by(state_name) %>%
  summarise_at(vars(area_km2,very_low_mf_own_mw:mod_sf_rent_mw),sum,na.rm=TRUE)
#View(capacity_potential_lmi_state)
str(capacity_potential_lmi_state)
```

#Scatter plot
```{r}
pairs(scale(capacity_potential_lmi_state[,3:14]))
```

#States names 
```{r}
num_vabs_cp_lmi_state <- as.data.frame(capacity_potential_lmi_state[,3:14])
rownames(num_vabs_cp_lmi_state) <- abbreviate(capacity_potential_lmi_state$state_name)
head(num_vabs_cp_lmi_state)
```


#Standardize
```{r}
num_vabs_cp_lmi_state.s <- normalize(num_vabs_cp_lmi_state, method = "standardize", range = c(0, 1), margin = 2, on.constant = "quiet")
head(num_vabs_cp_lmi_state.s)
```

#Correlation matrix
```{r}
corr_mat = cor(num_vabs_cp_lmi_state.s)
corrplot(corr_mat)
```

#Assumption: dataset is multivariate normality.

# Principal Component Analysis - States level
```{r}
lmi_pca_state <- princomp(num_vabs_cp_lmi_state.s, cor=T)
summary(lmi_pca_state,loadings=T)
```

#PC %Variance 
```{r}
fviz_screeplot(lmi_pca_state, addlabels = TRUE, ylim = c(0, 100))
```

#Biplot
```{r,dpi=1200,fig.height = 6, fig.width = 6, fig.align = "center"}
biplot(lmi_pca_state,cex=0.9,col=c('black','red'))
```


*Sun Belt states*
```{r}
sunbelt <- c('Florida','Georgia','South Carolina','Alabama','Mississippi', 'Louisiana','Texas','New Mexico','Arizona','Nevada','California')
capacity_potential_lmi_state$sunbelt <- 'Other'
capacity_potential_lmi_state$sunbelt[capacity_potential_lmi_state$state_name %in% sunbelt] <- 'SunBelt'
```

```{r}
num_vabs_cp_lmi_state$sunbelt <- capacity_potential_lmi_state$sunbelt
```


```{r,dpi=1200,fig.height = 6, fig.width = 6, fig.align = "center"}
autoplot(princomp(num_vabs_cp_lmi_state[,-13],cor=T),loadings=TRUE,loadings.label=TRUE,loadings.label.size=4,loadings.label.colour='black',data=num_vabs_cp_lmi_state,colour='sunbelt',loadings.colour='black',size=5)+
  theme(legend.position="bottom",legend.title=element_text(size=15),legend.text=element_text(size=14),axis.text=element_text(size=14),axis.title=element_text(size=14))

```


# Map PC1 scores, PC2 scores
```{r}
potential_state <- data.frame(state = capacity_potential_lmi_state$state_name, scores =lmi_pca_state$scores[,1])
rownames(potential_state) <- NULL
```

```{r,dpi=1500}
plot_usmap(data = potential_state, values = "scores", lines = "black") + 
  scale_fill_continuous(name = "PC1 Scores", label = scales::comma) + 
  theme(legend.position = "right")
```

#Inference PC1: 

PC1 shows the overall potential of solar rooftop energy capacity in different states for LMI households. It explains nearly 88% of the variance of original dataset. 

Based on PC1 scores, it would be obeserved that as expected, California, Texas and Florida have most extremely potential in solar rooftop energy capacity compared to other states. 

It is a little suprised that some north states, eg: New York, Pennsylvania, Massachusetts, New Jersey,... have considerable potential capacity for solar rooftop energy, even higher compared to other sunbelt states for LMI households.    


```{r}
potential_state_PC2 <- data.frame(state = capacity_potential_lmi_state$state_name, scores =lmi_pca_state$scores[,2])
rownames(potential_state_PC2) <- NULL
head(potential_state_PC2)
```

```{r,dpi=1500}
plot_usmap(data = potential_state_PC2, values = "scores", lines = "black") + 
  scale_fill_continuous(name = "PC2 Scores", label = scales::comma) + 
  theme(legend.position = "right")
```

#Inference PC2: 

It explains 10% of dataset variance

There is a contrast in solar rooftop potential capacity in 2 groups: multifamily-own and single-family-rent for LMI households.

For Texas and California, there is a higher focus of solar rooftop capacity potential in single-family-rent. 

For north states, eg: New York, New Jersey,... and Florida, there is a higher focus of solar rooftop capacity potential in multifamily-own. *WHY?*

##----------------------------------------------------------------------#

# Clusters Analysis  

## States level - SUM

**Hierarchical Clustering**

*overall sum capacity*
```{r}
dist_state <- dist(num_vabs_cp_lmi_state.s)
```

#Average linkage
```{r}
hc.s <- hclust(dist_state,'average')
plot(rev(hc.s$height))
```

#5 clusters
```{r}
ct.a <- cutree(hc.s,5)

```


#Complete linkage
```{r}
hc.s <- hclust(dist_state,'complete')
plot(rev(hc.s$height))
```


#single linkage
```{r}
hc.s <- hclust(dist_state,'single')
plot(rev(hc.s$height))
```


**Model-based clusering**
```{r}
mclust_state <- Mclust(num_vabs_cp_lmi_state[,1:12])
```


```{r}
summary(mclust_state)
```

```{r}
clust.data = cbind(mclust_state$classification,mclust_state$uncertainty)
clust.data[order(mclust_state$uncertainty,decreasing=TRUE),]
```

Plot of the the "uncertainty" plot for variables (SER, SPS) is:
```{r}
plot(mclust_state,what='uncertainty',dimens=c(6,8))
text(mclust_state$data[,c(6,8)],labels=rownames(num_vabs_cp_lmi_state),cex=0.6)
```

```{r}
subset(capacity_potential_lmi_state$state_name,mclust_state$classification == 1)
```

```{r}
subset(capacity_potential_lmi_state$state_name,mclust_state$classification == 2)
```

```{r}
subset(capacity_potential_lmi_state$state_name,mclust_state$classification == 3)
```


```{r}
plot(mclust_state,what='BIC')
```


## Clustering using Mean aggregation at State level

#PurposeAverage solar rooftop energy capacity per county for each state

#Reason: reduce the effect of size of large states eg: Cali, Texas,... to observe more the effect of other factors eg: housing type, tenure, income,...

#Assumption: county sample size for each state are approximately same

```{r}
capacity_potential_lmi_state_mn <- capacity_potential_lmi_clean %>%
  group_by(state_name) %>%
  summarise_at(vars(very_low_mf_own_mw:mod_sf_rent_mw),mean,na.rm=TRUE)
#View(capacity_potential_lmi_state)
str(capacity_potential_lmi_state_mn)
```

```{r}
mean_num_lmi_state <- capacity_potential_lmi_state_mn[,2:13]
```

```{r}
mean_num_lmi_state_s <- scale(mean_num_lmi_state)
```

```{r}
corr_mat <- round(cor(mean_num_lmi_state_s),2)
corrplot(corr_mat)
```

#Principal Component Analysis
```{r}
lmi_mn_pc <- princomp(mean_num_lmi_state_s,cor=T)
summary(lmi_mn_pc, loading=T)
```

PC1: the overall average capacity per county for each states 

PC2: the contrast between mf_own and sf_rent, the higher score, the higher from mf_own

PC3: the contrast between Risk-free (mf_own, sf_rent)v Risk-seeker(mf_rent,sf_own), the higher the less risk

```{r}
cbind(lmi_mn_pc$scores[,1],capacity_potential_lmi_state_mn$state_name)
```

```{r}
plot.wgss = function(mydata, maxc) {
  wss = numeric(maxc)
  for (i in 1:maxc) 
    wss[i] = kmeans(mydata, centers=i, nstart = 10)$tot.withinss 
  plot(1:maxc, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares", main="Scree Plot") 
}
```

```{r,dpi=1200}
plot.wgss(mean_num_lmi_state_s,20)
abline(v=3,col='red',lty=3)
```

```{r}
fviz_nbclust(mean_num_lmi_state_s, kmeans, method = "wss") +
    geom_vline(xintercept = 3, linetype = 2,color='red')
```


```{r}
km <- kmeans(mean_num_lmi_state_s, centers = 3, nstart = 30) 
table(km$cluster)
```

```{r}
rownames(mean_num_lmi_state_s) <- capacity_potential_lmi_state_mn$state_name
```


```{r,dpi=1500}
fviz_cluster(km, data = mean_num_lmi_state_s,
            ellipse.type = "norm", repel = FALSE, labelsize = 13
             )
```

```{r}
km$cluster
```


#3D plotting
```{r}
#plot3d(lmi_mn_pc$scores[,1:3],type='s', size=1,col=km$cluster)
#text3d(lmi_mn_pc$scores[,1:3],text=abbreviate(capacity_potential_lmi_state_mn$state_name),col=km$cluster)
```

```{r}
lmi_scores <- as.tibble(lmi_mn_pc$scores)
lmi_scores$Cluster <- as.factor(km$cluster)
lmi_scores$States <- capacity_potential_lmi_state_mn$state_name
```

```{r}
#lmi_scores
```

```{r,dpi=1200}
#plot(lmi_mn_pc$scores[,1:2],fill=km$cluster)
ggplot(lmi_scores,aes(Comp.1,Comp.2,color=Cluster))+
  geom_point(size=2)+
  geom_text(aes(label=States))
  
```

```{r,dpi=1200}
ggplot(lmi_scores,aes(Comp.1,Comp.3,color=Cluster))+
  geom_point(size=2)+
  geom_text(aes(label=States))

```


```{r,dpi=1200}
ggplot(lmi_scores,aes(Comp.2,Comp.3,color=Cluster))+
  geom_point(size=2)+
  geom_text(aes(label=States))
```

```{r}
biplot(lmi_mn_pc)
```

```{r setup}
#library(knitr)
#library(rgl)
#knit_hooks$set(webgl = hook_webgl)
```

```{r testgl, webgl=TRUE}
#x <- sort(rnorm(1000))
#y <- rnorm(1000)
#z <- rnorm(1000) + atan2(x,y)
#plot3d(x, y, z, col=rainbow(1000))
```


#Model-based clustering
```{r}
mc <- Mclust(mean_num_lmi_state_s)
```

```{r}
summary(mc)
```



##------------------------------------------------------------------------------##

# EFA & CCA

##Type of buildings and its tenure which has high potential rooftop solar energy potential.
```{r}
#str(num_vabs_cp_lmi)
```

#Factor analysis with cutoff = 0.5 for 2 factors 
```{r}
lmi_fa <- factanal(num_vabs_cp_lmi,factors=2)
lmi_fa
print(lmi_fa, cut=0.5)
```

```{r}
loadings_lmifa <- lmi_fa$loadings
loadings_lmifa
(loadings_lmifa_df <- as.data.frame(as.table(loadings_lmifa)))

ggplot(loadings_lmifa_df,aes(Var1,Freq,fill=Freq))+
  geom_bar(stat='identity')+
  coord_flip()+
  facet_wrap(~Var2,nrow=1)+
  scale_fill_gradient2(name='loading',midpoint =0.50 ,guide='colourbar')+
  theme_bw(base_size=10)+
  xlab('Variables')+
  ylab('Loading Strength')+
  theme(axis.title=element_text(size=13),axis.text=element_text(size=13),
        strip.text.x=element_text(size=13))+
  geom_hline(yintercept=0.5,color='red')
```

```{r}
f.loadings <- lmi_fa$loadings[,1:2]
corHat <- f.loadings%*%t(f.loadings)+diag(lmi_fa$uniquenesses)
sqrt(mean((cor(num_vabs_cp_lmi)-corHat)^2))

```


## 2 Factors: Community sharing and Personal use 
```{r}
bt_model <- specifyModel(text= 'Personal -> very_low_sf_own_mw, lambda1, NA
                                Personal -> very_low_sf_rent_mw, lambda2, NA
                                Personal -> low_sf_own_mw, lambda3, NA
                                Personal -> low_sf_rent_mw, lambda4, NA
                                Personal -> mod_sf_own_mw, lambda5, NA
                                Personal -> mod_sf_rent_mw, lambda6, NA
                                Personal -> very_low_mf_rent_mw, lambda7, NA
                                Personal -> low_mf_rent_mw, lambda8, NA
                                Personal -> mod_mf_rent_mw, lambda9, NA
                                Community -> very_low_mf_own_mw, lambda10, NA
                                Community -> low_mf_own_mw, lambda11, NA
                                Community -> mod_mf_own_mw, lambda12, NA
                                Personal <-> Community, rho, NA
                                very_low_sf_own_mw <-> very_low_sf_own_mw, theta1, NA
                                very_low_sf_rent_mw <-> very_low_sf_rent_mw, theta2, NA
                                low_sf_own_mw <-> low_sf_own_mw, theta3, NA
                                low_sf_rent_mw <-> low_sf_rent_mw, theta4, NA
                                mod_sf_own_mw <-> mod_sf_own_mw, theta5, NA
                                mod_sf_rent_mw <-> mod_sf_rent_mw, theta6, NA
                                very_low_mf_own_mw <-> very_low_mf_own_mw, theta7, NA
                                very_low_mf_rent_mw <-> very_low_mf_rent_mw, theta8, NA
                                low_mf_own_mw <-> low_mf_own_mw, theta9, NA
                                low_mf_rent_mw <-> low_mf_rent_mw, theta10, NA
                                mod_mf_own_mw <-> mod_mf_own_mw, theta11, NA
                                mod_mf_rent_mw <-> mod_mf_rent_mw, theta12, NA
                                Personal <-> Personal, NA, 1
                                Community <-> Community, NA, 1')
```


```{r}
bt_sem2 <- sem(bt_model,cor(num_vabs_cp_lmi), nrow(num_vabs_cp_lmi))
summary(bt_sem2)
```


```{r,dpi=1200}
semPaths(bt_sem2,rotation=2,'est',edge.label.cex = 1.2)
```

SRMR, GFI, and AGFI:
```{r}
options(fit.indices = c('GFI','AGFI','SRMR'))
criteria = summary(bt_sem2)
```

```{r}
criteria$SRMR
```

```{r}
criteria$GFI
```

```{r}
criteria$AGFI
```
